### 2016-08-23 syntax.tex
syntax

\heading{Preliminaries. } When~$\univ$ is a set and $n>0$ is an
integer, we let $[\univ]^n$ denote the set of all size-$n$ subsets
\jnote{Multisets?} of~$\univ$. We write $x \getsr \univ$ to denote
sampling an element from~$\univ$ and assigning this to~$x$, and we
extend this notation to randomized algorithms.  When~$X$ is a string
over some alphabet~$\Sigma$, we write $|X|$ for the bitlength
of~$X$, relative to some fixed encoding.  When~$X$ is a multiset, we
overload the notation so that~$|X|$ is the number of elements in~$X$
(with mutiplicity).

@@

\heading{Set-multiplicity Data Structures. } Our particular focus
will be on data structures that support multiplicity queries, of
which set-membership queries are a special case.
\begin{definition} A \emph{set-multiplicity data structure} is a data
structure with $D=\mathbb{N}$ \jnote{Notation. (Anyway, I don't
think we need an explicit symbol for the range of queries.)} and
where $\calQ=\{q_x \,|\, x \in \univ\}$ where $q_x(S)$ is defined to
be the multiplicity of~$x$ in the multiset~$S$. \hfill\dqed
\end{definition}
%
%\todo{Readdress issue of representation size.  A general statement might be something like `` The \emph{size} of a set-representation is the maximum (bit)length of any~$M$ that can be output by $\Rep$.''  But all of the natural constructions will have representation whose size will not depend on the coins of $\Rep$, i.e. they depend on $|S|$, number of hash functions, desired error rate. }

@@

%\heading{Bloom Filters and Related Data Structures. }
It is not hard to see Bloom filters are one possible instantiation a
set-multiplicity data structure. Fix an $n \geq 0$, let
$\mathcal{S}=[\univ]^n$, and let $\Sigma=\Gamma=\bits$. Define the
queries so that $q_x(S)=1$ if $x \in S$, and 0 otherwise.  Define
algorithm~$\Rep$, with functions $k=k(|S|), m=m(|S|)$ and a
description of a hash function family~$\mathcal{H}=\{h \colon \univ
\to \{1,2,\ldots,m\}\}$ hardwired into it, to sample hash functions
$h_1, \ldots, h_k$ from~$\mathcal{H}$ and then compute an $m$-bit
array in the standard way.  It sets $\privaux=\pubaux = (h_1,
\ldots, h_k)$ and~$M$ to be the bit array, and returns these.  The
algorithm~$\Qry$ is defined so that, on input $(M,\privaux,q_x)$, it
returns the AND of the values (either 0 or 1) held at
$M[h_1(x)],\ldots,M[h_k(x)]$.  We note that, while our error
probability and error rate do not distinguish among ``types'' of
errors, they also do not preclude instantiations (like the classical
Bloom filter) that admit only one type of error, e.g.
false-positives.

>>>>

\heading{Preliminaries. } When~$\univ$ is a set and $n>0$ is an
integer, we let $[\univ]^n$ denote the set of all size-$n$ subsets
%\jnote{Multisets?} 
of~$\univ$. We write $x \getsr \univ$ to denote
sampling an element from~$\univ$ and assigning this to~$x$, and we
extend this notation to randomized algorithms.  When~$X$ is a string
over some alphabet~$\Sigma$, we write $|X|$ for the bitlength
of~$X$, relative to some fixed encoding.  When~$X$ is a multiset, we
overload the notation so that~$|X|$ is the number of elements in~$X$
(with mutiplicity).

@@

\heading{Set-multiplicity Data Structures. } Our particular focus
will be on data structures that support multiplicity queries, of
which set-membership queries are a special case.
\begin{definition} A \emph{set-multiplicity data structure} is a data
structure with $\Gamma=\mathbb{N}$ and
where $\calQ=\{q_x \,|\, x \in \univ\}$ where $q_x(S)$ is defined to
be the multiplicity of~$x$ in the multiset~$S$. \hfill\dqed
\end{definition}
%
%\todo{Readdress issue of representation size.  A general statement might be something like `` The \emph{size} of a set-representation is the maximum (bit)length of any~$M$ that can be output by $\Rep$.''  But all of the natural constructions will have representation whose size will not depend on the coins of $\Rep$, i.e. they depend on $|S|$, number of hash functions, desired error rate. }

@@

%\heading{Bloom Filters and Related Data Structures. }
It is not hard to see Bloom filters are one possible instantiation a
set-multiplicity data structure.  Fix $\Sigma=\Gamma=\bits$.  Define
algorithm~$\Rep$, with functions $k=k(|S|), m=m(|S|)$ and a
description of a hash function family~$\mathcal{H}=\{h \colon \univ
\to \{1,2,\ldots,m\}\}$ hardwired into it, to sample hash functions
$h_1, \ldots, h_k$ from~$\mathcal{H}$ and then compute an $m$-bit
array in the standard way.  It sets $\privaux=\pubaux = (h_1,
\ldots, h_k)$ and~$M$ to be the bit array, and returns these.  The
algorithm~$\Qry$ is defined so that, on input $(M,\privaux,q_x)$, it
returns the AND of the values (either 0 or 1) held at
$M[h_1(x)],\ldots,M[h_k(x)]$.  We note that, while our error
probability and error rate do not distinguish among ``types'' of
errors, they also do not preclude instantiations (like the classical
Bloom filter) that admit only one type of error, e.g.
false-positives.
